### 2018-05-24 abstract.tex
initial port from grove-paper repo

% !TEX root = palettes-paper.tex

\begin{abstract}

Palettes... \cite{popl-paper}

\end{abstract}