### 2019-02-26 abstract.tex
-nested

% !TEX root = palettes-paper.tex

%% no citations in Abstract
%%
%% \cite{popl-paper}

\begin{abstract}
%
\vspace{2.00in} %% TODO leave room for authors/copyright
%
We present \emph{palette expressions}, an approach to programming custom
graphical user interfaces (GUIs) that are used to build expressions---to ``fill
holes''---within an incomplete program.
%
Palette expressions serve as an alternative and complementary code editing
mechanism to traditional text-editing.
%
Compared to other ``hybrid'' text-and-GUI code editors, palette expressions
provide several unique capabilities.
%
First, a palette persists even after user interactions produce a complete
expression that fills the hole, so that the programmer can return to the GUI to
make subsequent changes.
%
Second, palettes for smaller expression types can appear nested within those for
larger expressions, allowing palettes to be composed in complex ways.
%
Third, palettes have access to the execution environment---even when the overall
program is not yet complete---and can, thus, provide live feedback to the
programmer while working to fill an incomplete expression.

This paper investigates the design and implementation of palettes in several
steps.
%
First, we propose a formal framework for palettes within Hazelnut, a
foundational semantics for live programming that provides continuous code editor
services through a combination of structured edits and dynamic evaluation of
incomplete programs.
%
Second, to demonstrate the expressiveness of our framework, we describe several
example palettes, many of which are informed by a prior study that identified
desirable use cases for hybrid editors.
%
Lastly, we provide a prototype implementation of palettes within Hazel, a
programming environment based on the Hazelnut theory extended with palettes.
%
Based on our experience, we believe palettes are an extensible and expressive
approach for mixing text- and GUI-based code editing.

\end{abstract}