### 2019-02-26 evaluation.tex
rough outline notes from Friday

\section{Evaluation}

\subsection{Implementation}

implementation in \Hazel{}: palette macros; structure edit actions; expression formula bar; inline layout

implementation in \sns{}: palette functions; text editor; pop-up menu

\subsection{Examples}

\rkc{old list of examples below:}

\subsubsection{Boolean}
Checkbox

\subsubsection{Matrix}
Simple example

\subsubsection{Grade Cutoffs}
Uses liveness more obviously

\subsubsection{Table}
Uses type reflection

\subsubsection{Pixel Art}
Cool example

\subsubsection{Regex}
Similar to Graphite paper, can cite the empirical study we did there

\subsubsection{Forms}
Show off composition + mention full-screening stuff

\subsubsection{Equation Editor, Judgement Editor and Category Diagrams}
Things that PL people like

\subsubsection{TikZ diagrams}
...