### 2019-02-26 overview.tex

\section{Overview Examples: Using Palettes}
\label{sec:overview}

\begin{figure*}

\includegraphics[scale=0.50]{images/MrSmileyFace.png}

\caption{Example Palettes in Hazel.}
\end{figure*}

\begin{itemize}

\item figure with 2/3 Hazel palettes:

  \begin{enumerate}
    \item table (synthetic)
    \item grade cutoffs (analytic)
    \item color? (analytic)
  \end{enumerate}

\item emphasize persistence

\item emphasize composition

\item emphasize liveness

\item mention analytic (annotated) vs. synthetic (unannotated) palettes

  \begin{itemize}
    \item stick to palette macros for now
    \item introduce palette functions in ``Overview Examples: Defining Palettes''
          or ``Formal System''
  \end{itemize}

\end{itemize}