### 2020-05-24 discussion.tex
separated everything into separate files

\section{Discussion and Conclusion}\label{sec:discussion}\label{sec:conclusion}
Future work:
\begin{itemize}
  \item pattern matching
  \item type splices
  \item module splices
  \item explicitly make bindings available in splices
  \item better UI for closure provenance / connect with control flow / call stack better
  \item deriving livelits from type definitions
  \item bidirectional evaluation ala SnS
  \item integration with structure editing more cleanly
  \item \dots
  \item collaborative editing?
  \item side effects?
  \item full screening livelits (or compositions thereof) as a way to create end-user workflows
  \item big vision: authoring environment based on livelits
\end{itemize}