### 2020-05-24 implementation.tex

\section{Implementation}\label{sec:implementation}
What should we talk about here? Layout of livelits?