### 2020-05-24 livelit-definitions.tex

\section{Livelit Definitions}\label{sec:livelit-definitions}
Which livelit implementation do we want to show? Maybe grade cutoffs? Slider as simple intro + color chooser?

Talk about each component of a livelit definition:
\begin{itemize}
  \item livelit name (talk about binding structure)
  \item expansion type
  \item model type (must be serializable)
  \item action type (mention integration into Hazel's undo system)
  \item initial model
  \item update function (monadic actions; talk about color palette update; SpliceRefs)
  \item view function (monadic actions; talk about GUI stuff? tab order? dimensions? mostly defer to Elm for details)
  \item expand function (should we call it fill?)
\end{itemize}

Hygiene from a provider's perspective -- have to use parameters to explicitly capture definition site bindings.