### 2020-07-03 livelit-calculus.tex

\section{A Typed Livelit Calculus}\label{sec:livelit-calculus}
Should we talk about anything other than expansion + hygiene, i.e. parameterization, closure collection, closure resumption?
None of that is mechanized, but that may be okay. Closure collection and resumption seem to be worth formalizing.

Note: probably just going to omit the whole livelit function thing

Subsection about Agda mechanization (Ian + Nick?)