### 2020-11-18 popl21/popl21-rebuttal.txt
moved popl21 submission into a directory for archival

Thank you for your reviews!

## Review A

### Corrections

The quoted sentence was part of our discussion of _future_ work. It was not written as motivation for this paper.

Livelits can generate expressions, not just values. This includes expressions that have non-trivial computational behavior. 
For example, Fig 3 demonstrates image transformations.

## Review B

### Comments

Although the five examples in the paper are non-trivial, we agree that additional screenshots might help readers build 
better intuitions for the expressive power of livelits. Consequently, we would be happy to include 4-6 additional 
screenshots of novel livelits in the form of an "example gallery" in our final revision. None of these introduce
new technical content into the paper.

Our prior work on Graphite included several empirical studies which are just as relevant to livelits.
The present paper can be read as a technical complement to that qualitative/empirical paper [Omar et al, 2012].

The Hazel livelits implementation as described in the paper was functional at the time of submission
(in the livelits branch on Github). It will be sent for artifact evaluation and released publicly with the 
next major Hazel release before POPL. The public will be able to play with livelits in their browser. We are 
working on a number of ``quality of life'' improvements on an ongoing basis, but these do not change the 
content or contributions of the paper.

## Review C

### Corrections

MPS and Barista (via Citrus) were discussed in detail in the paper. These systems, as well as the others that 
the review mentions, fail to achieve multiple of the design goals from Sec. 1 (discussed on line 921). We
will include missing citations in the final revision, thank you.

Livelits could be defined using other livelits when using the definition mechanism defined in Sec. 3. It is not the case that ``the definitions for livelits are textual'' in this setting. Indeed, related work on literal macros supports using literal macros in literal macro definitions [Omar et al, 2018] and the same mechanisms could be deployed here (once Hazel has a proper module / package management system).

We believe pure functional programming in languages like Haskell and Elm is practical. 
That said, it is not true that livelits are limited to the pure functional setting. 
The issue is simply that running a program with side effects
during editing might not be _desirable_ in some settings, but that choice is up to the programmer.
For example, live programming in Jupyter lab notebooks often involves side effects 
occurring in arbitrary order depending on user interactions and programmers by and large accept this. 
Livelits could be added to Jupyter-like systems such that they always use this ambient state.
The issue could also be avoided by using algebraic effects, so that destructive effects could be ``shimmed'' 
during development. These considerations are not livelit specific -- they are inherited from the work 
in [Omar et al., POPL 2019] and indeed live programming itself is fundamentally in tension with 
side effects. We will include a more substantial discussion of this point in the final revision.

We do discuss livelit persistence in Sec. 3.1. Only the model is serialized. In our case, we use a generic S-expression
textual syntax for Hazel expressions, but this is an incidental implementation detail (Hazel programs are not
intended to be edited using text editors). We do discuss a proof-of-concept text editor with support for 
livelits in Sec. 5.2.

### Comments

Pattern livelits would be a straightforward extension of our system. The necessary theory for pattern macros
was developed in [Omar et al, 2018].

Declaration livelits can be subsumed by module livelits, which can in turn be subsumed by expression livelits
in a language that has first class modules. This was also briefly discussed in [Omar et al, 2018].

While our formalism of expansion does build on our prior macro system formalism, we see this as a strength
of our paper. Our treatment of splices is distinct, as is our treatment of hygiene. Closure collection is entirely
new. The purpose of the formalism was to make the mechanism precise and reduce it to its essential form.

Hazel functions that use GUI libraries must ultimately execute by extraction to OCaml, so the fact that 
we implemented our GUI code directly in OCaml was merely a matter of expediency. The Hazel language is still
evolving (in large part due to contributors who are not involved with this paper) and there is not yet a 
full-scale GUI widgets library. We plan to "self host" in the coming months before POPL.

The interface for selecting between closures is the same as the interface demonstrated in [Omar et al, POPL 2019]. We will include a screenshot in the final revision.

The capture list in Fig. 4 is empty because generated expansions do not use any libraries or definition site values
(they are simple tuples of numbers in this case). See [Omar et al, ICFP 2018] for a discussion on explicit capture.

## Review D

### Corrections

The ipywidgets library only allows simple non-compositional widgets (no splices inside) and these widgets do not have access to the surrounding 
closures or state so there is no support for liveness. 
None of the central examples in our paper could be implemented as ipywidgets. We will, however, make sure to cite this work in the final revisions.

### Comments

See our response to Review B regarding user studies.

Typed holes are central in that they enable closure collection, which is what enables liveness.

The core calculus includes sums, products, and recursive types because these are minimal type theoretic constructs sufficient to achieve
self-hosting of syntax. Alternative encoding schemes could work just as well (e.g. Church encodings in System F) but are more unwieldy.
Self-hosting is necessary because the expand function returns an encoding of syntax, which needs to be decoded (Sec. 4.2.1).

Splices are simply expressions that appear inside (i.e. are notionally spliced into) livelit GUIs.

We will cite Lerner 2020, which has related motivations, thank you.

We will use a different connective for expanding and elaborating in the final version, thank you.