### 2020-11-19 pldi21/abstract.tex
further tightening of abstract  typo fix

% !TEX root = livelits-paper.tex

%% no citations in Abstract
%%
%% \cite{popl-paper}

\begin{abstract}
    Although text editing is powerful, 
    some types of expressions are more naturally represented and manipulated
    graphically.
    % Examples include expressions that compute 
    % colors,
    % animation parameters,
    % musical sequences,
    % tabular data, 
    % plot parameters,
    % GUI widgets, 
    % and various other domain-specific data structures.
    This paper introduces \emph{live literals}, or \emph{livelits}, 
    which allow
    clients to fill holes of these types  
    by directly manipulating a     
    provider-defined GUI embedded persistently into otherwise 
    symbolic code. 
    Uniquely, livelits are compositional: the GUI can itself contain
    spliced expressions, which have full editor support and can in turn contain other livelits. 
    Splices are typed and the system 
    ensures that livelits treat splices hygienically. 
    Livelits are also uniquely live: 
    they can offer {immediate feedback} about the run-time implications 
    of the client's choices 
    even when splices mention bound variables, 
    because the system continuously gathers closures associated with the 
    hole that the livelit is tasked with filling.
    This paper introduces livelits with case studies that exercise these novel capabilities. 
    We implement livelits in Hazel, 
    a live programming environment able to 
    typecheck and run programs with holes. 
    We then define a typed lambda calculus that captures the essence of livelits as 
    live graphical macros.     
    The macro expansion process has been mechanized in Agda.
\end{abstract}