### 2021-04-13 pldi21/livelit-definitions.tex
updates in respond to shepherd

\section{Livelit Definitions}\label{sec:livelit-definitions}

\begin{figure}
\begin{lstlisting}[xleftmargin=0.2cm]
type Color = (.r Int, .g Int, .b Int, .a Int)
livelit $color at Color {
  type Model = (.r SpliceRef, .g SpliceRef, 
                .b SpliceRef, .a SpliceRef)

  context { }

  let init : UpdateCmd(Model) = do 
    r <- new_splice(`Int`, Some(`0`))
    g <- new_splice(`Int`, Some(`0`))
    b <- new_splice(`Int`, Some(`0`))
    a <- new_splice(`Int`, Some(`100`))
    return (r, g, b, a)

  type Action = 
  | ClickOn(Color)

  let view : Model -> ViewCmd(Html(Action)) = 
    fun model -> do 
      (* determine a color to display *)
      r_res <- eval_splice(model.r)
      g_res <- eval_splice(model.g)
      b_res <- eval_splice(model.b)
      a_res <- eval_splice(model.a)
      let cur_color : Color = 
        case (r_res, g_res, b_res, a_res) 
        | (Some(Val(IntLit(r))), 
           Some(Val(IntLit(g))), 
           Some(Val(IntLit(b))), 
           Some(Val(IntLit(a)))) -> 
             Some((r, g, b, a))
        | _ -> 
          (* indeterminate color shown as X *)
          None 
      in

      (* generate splice editors *)
      let size = FixedWidth(20) in 
      r_editor <- editor(model.r, size)
      g_editor <- editor(model.g, size)
      b_editor <- editor(model.b, size)
      a_editor <- editor(model.a, size)

      (* ... now we can render the UI ... *)

  let update : 
      Model -> Action -> UpdateCmd(Model) =
    fun model (ClickOn c) -> do 
      set_splice(model.r, IntLit(c.r))
      set_splice(model.g, IntLit(c.g))
      set_splice(model.b, IntLit(c.b))
      set_splice(model.a, IntLit(c.a))
      return model

  let expand : Model -> (Exp, List(SpliceRef)) = 
    fun model -> (`fun r g b a -> (r, g, b, a)`, 
      [model.r, model.g, model.b, model.a])
}
\end{lstlisting}
\vspace{-8px}
\caption{Livelit Implementation}
\label{fig:color-impl}
\end{figure}

\noindent
We will now take the perspective of a livelit provider.
Fig.~\ref{fig:color-impl}
defines \li{\$color} from Fig.~\ref{fig:color}, 
which is our prototypic example of a livelit 
definition. 
We omit certain  
incidental details and use unimplemented syntactic sugar, including
Haskell-style \li{do} notation \cite{marlow2010haskell} and quasiquotation \cite{bawden1999quasiquotation}, for presentation.

Livelit definitions are scoped and packaged like 
any other definition. Each \emph{definition} consists of a \emph{declaration} and an \emph{implementation}. 
In Hazel, a ``template'' declaration and implementation is generated 
as soon as the user types ``\li{livelit\ }''.

\subsection{Livelit Declarations} 
Line 2 of Fig.~\ref{fig:color-impl} is \li{\$color}'s declaration,
which defines its name and its {expansion type}, \li{Color}, defined on Line 1. 
Both are required. 
Livelit parameters can also optionally appear here as shown in the declarations in Sec.~\ref{sec:expansion-typing}. 
The declaration is part of the client interface, 
as discussed in Sec.~\ref{sec:expansion-typing}-\ref{sec:parameterization}.

\subsection{Livelit Implementations}
The curly braces delimit the livelit's implementation, which is not intended to be seen 
by clients. Each livelit implementation must define types \li{Model} and \li{Action}, 
values \li{init}, \li{view}, \li{update}, and \li{expand} with the signatures specified in Fig.~\ref{fig:color}, 
and a \li{context}, which is a listing of definition-site bindings that the livelit can make use of as described below. All of these are included in the generated template.
 These constitute a variation on the pure functional model-view-update
architecture popularized by Elm \cite{ElmArchitecture}. We add a fourth component,
expansion generation. In addition, we use a monadic framework (\emph{a la} Haskell \cite{marlow2010haskell}) to provide a pure   
 interface between the livelit and the editor: monadic commands, like \li{new_splice} discussed below, are executed by the editor,
and \li{do} notation provides syntactic sugar for monadic bind.

\subsubsection{Model}\label{sec:model}
The state of a livelit's GUI is determined by its model value. 
Line 3 of Fig.~\ref{fig:color-impl} specifies the corresponding \emph{model type},
here a labeled 4-tuple of \emph{splice references}, one for each of the four splices
that appear in the GUI in Fig.~\ref{fig:color}.
The model is how the GUI state is persisted in the syntax tree, so 
 the system requires that the model type supports automatic serialization
 (so functions cannot appear in models).

The \li{init} value on Line 8 determines the value of the model
when the livelit is first invoked in the editor.
It is a command in the \li{UpdateCmd} monad, further discussed in Sec.~\ref{sec:def-update}, that returns the 
initial model value after generating four new splices 
using the \li{new_splice} command:
\begin{lstlisting}[numbers=none]
  new_splice : (Typ, Maybe(Exp)) 
               -> UpdateCmd(SpliceRef)
\end{lstlisting}
This command creates a splice
of the given type and, optionally, its initial contents. 
It returns a splice reference, which uniquely identifies that splice.
In this section, bolded types  
are defined in the standard library. The \li{Typ} and \li{Exp} types 
encode the syntax of Hazel's types and expressions 
and we use quasiquotation, e.g. \li{`0`}, as the introduction forms \cite{bawden1999quasiquotation}. 
% If no initial expression is provided, the splice starts as  
% an empty hole.

To ensure \emph{context independence}, 
the system checks that the splice type and initial content are valid 
assuming only the parameters and explicitly specified \li{context} on Line 6.
Here, the context is empty because \li{Int} is a built in type
and the initial expressions are integer literals. 
We use an explicit context, rather than implicitly capturing all bindings 
at the definition site, 
to ensure that private bindings are not unintentionally  
leaked to clients \cite{TLMs}.
% Formally, captured variables can be understood as parameters that have been 
% immediately partially applied.

\subsubsection{Action}
Line 15 defines the \li{Action} type for the \li{\$color} livelit, which 
specifies a single user-initiated action: clicking on a color using
the right half of Fig.~\ref{fig:color}. Actions are emitted
from event handlers (e.g. click handlers) defined in the computed \li{view}, Sec.~\ref{sec:def-view}, 
and actions are consumed by the \li{update} function, Sec. \ref{sec:def-update}, causing a change to the model.

\subsubsection{View}\label{sec:def-view}
The \li{view} function computes the view given the model and access to the commands in 
the \li{ViewCmd} monad (which is distinct from the \li{UpdateCmd} monad). The computed view is a value of type \li{Html(Action)}.  
This type  provides a simple immutable
encoding of an HTML element, where the type parameter is the type of actions that 
are emitted by event handlers that can be attached to elements, e.g.
\li{on_click} and so on. 
We elide the details of the particular user interface in Fig.~\ref{fig:color}, 
but note that livelit implementations can themselves invoke other  
livelits, e.g. the \li{view} function could use livelits for 
expressing user interface widgets and layouts (not shown). 
Instead, we focus on three mechanisms exposed by \li{ViewCmd}: live evaluation,
splice editors, and result rendering.

\paragraph{Live Evaluation}
\label{sec:live-evaluation-def}
As discussed in Sec.~\ref{sec:live-evaluation}, 
the view can depend on the result of evaluating a splice or a parameter
under the closure the client has selected.
(Parameters, not shown in this example but discussed in Sec.~\ref{sec:parameterization}, operate like splices and have type \li{SpliceRef}
within the livelit definition.)
The interface between the view and the live evaluator 
is via the 
following command:
% Actually, it can depend on an encoded function of 
% multiple splices as long as the function is well-typed using only the capture set, 
% under the closure that the client has selected using one of the following
% commands:
\begin{lstlisting}[numbers=none,xleftmargin=0cm]
eval_splice : SpliceRef -> ViewCmd(Maybe(Result))
\end{lstlisting}

%eval_fun : (Exp, List(SpliceRef)) -> ViewCmd(Maybe(Result))

The \li{None} case 
arises when evaluation is not possible, e.g. because no closures are collected 
or because no value has been collected for a variable used in the splice. 
In practice, there is an implicit hole at the end of each cell in Hazel, so 
livelits will have at least one closure containing results for at least 
the top-level bindings. Function argument values may not be available 
if that function has not been applied.
% because the free variables in the selected closure overlap with the free 
% variables in the provided expression. This occurs only when the user has selected
% a closure that appears under a binder in the evaluation result (e.g. because 
% the result is a lambda). 
% We discuss this further in Sec.~\ref{sec:calculus-closure-collection}.

If available, \li{Result} distinguishes two possibilities:
\begin{lstlisting}[numbers=none]
type Result = Val(Exp) | Indet(Exp)
\end{lstlisting}
The \li{Val} case arises when evaluation produces a value, whereas the 
\li{Indet} case arises when evaluation results in an indeterminate expression,
i.e. an expression that cannot be fully evaluated due to holes 
in critical positions \cite{HazelnutLive}.

Lines 26-34 determine a color to display in the color preview 
if all four splices evaluate to integers. Otherwise, there is not 
enough information to determine a color. The livelit 
indicates this situation by disabling the color preview.

Livelits can attempt to offer feedback even when the result is indeterminate,
because indeterminate expressions might nevertheless contain useful information.
For example, a livelit that previews a sequence of notes as audio might be able 
to handle a list of notes where certain notes are missing, i.e. holes, by 
playing silence.
This behavior is highly domain-specific, so each livelit provider must decide 
whether and how indeterminate results are supported.

\paragraph{Splice Editors}
The view includes an editing area for each splice. These editors must support all of 
Hazel's editing services. To support this, the \li{view} function
can request an editor with a given dimension (in character units) for a given splice:
\begin{lstlisting}[numbers=none]
editor : (SpliceRef, Dim) -> ViewCmd(Html(a))
\end{lstlisting}
The result is an opaque \li{Html(a)} value that the remainder of the function 
can place where needed. When the livelit is rendered, this part of the tree is 
under the control of Hazel. The \li{Dim} parameter currently supports only a fixed
character width, with overflow causing scrolling, but in the future we plan to offer 
to offer more flexible layout options.

\paragraph{Result Rendering}
Some livelit views needs to include 
a rendered evaluation result. For example, each of the cells in the \li{\$dataframe}
livelit in Fig.~\ref{fig:grading} show the evaluation result for the corresponding 
cell. Only the formula bar at the top is an editor. To support this, the \li{view}
function can use the \li{result_view} command, which mirrors the \li{editor} command:
\begin{lstlisting}[numbers=none]
result_view : (SpliceRef, Dim) 
              -> ViewCmd(Maybe(Html(a)))
\end{lstlisting}
\subsubsection{Update}\label{sec:def-update}
When the user triggers an event in a livelit view, it emits an \li{Action}.
The system responds by calling the \li{update} function to determine how 
this action should affect the model and, in some cases, the splices.

% We discussed in Sec.~\ref{sec:model} that the \li{UpdateCmd}
% allows new splices to be generated in response to user actions. 
% For example, the \li{\$dataframe} livelit uses this when new rows
% or columns are added.

In Fig.~\ref{fig:color-impl}, we see that the \li{\$color} livelit responds to 
the \li{ClickOnColor} action by invoking the \li{set_splice} command to overwrite 
the current splices with integer literals determined based on which color the user
clicked on:
\begin{lstlisting}[numbers=none]
  set_splice : (SpliceRef, Exp) -> UpdateCmd ()
\end{lstlisting}
As with \li{new_splice} described in Sec.~\ref{sec:model},
the system maintains context independence by checking the expression against the splice type 
and only allows use of the specified \li{context}.

When the model is updated, a new view is 
computed. The system then performs a diff between the old and new view in order to 
efficiently perform the necessary imperative updates to the editor's visual state.
Changes to splices can also cause the view to be recomputed, because the view might 
evaluate the splices. The \li{UpdateCmd} monad does not itself 
have the ability to request evaluation (\li{eval_splice}), because the model should not depend directly  
on which closure the user has selected. 
Of course, the \li{view} might emit 
result-dependent actions when appropriate.

\subsubsection{Expansion}
\label{sec:expansion}
The ultimate purpose of a livelit is to fill the hole where it appears by generating an expansion,
i.e. an expression of the expansion type, here \li{Color}.
The \li{expand} function determines the expansion given the model. 
The two are necessarily distinct: the model encodes the GUI state, 
and so usually tracks ephemeral widget state, like which tab is active in a tabbed GUI, 
as well as splice references, which refer only indirectly to the spliced expressions themselves, discussed below. 
This information is not needed at run-time as part of the expansion.
Only the model is persisted when a Hazel program is saved, because the expansion 
can be regenerated by calling \li{expand} whenever needed.

The expansion can include spliced expressions, 
but the system does not make these expressions  
available directly as values of type \li{Exp}. 
Instead, the expansion must treat splices parametrically. 
In particular, \li{expand} returns an encoded \emph{parameterized expansion}, of type \li{Exp}, 
paired with a list of \li{SpliceRef}s (which can come from 
both splices and parameters as discussed above).
The {parameterized expansion}
  is a function 
 (here curried)
 that takes an argument for each listed \li{SpliceRef}. 
 That argument is of the corresponding splice 
type, which was provided when the splice was initialized. 
 The return type of the parameterized expansion is the expansion type.
 So here, the parameterized expansion for \li{\$color} must be a function of type 
 \li{Int -> Int -> Int -> Int -> Color} because the splices 
 in the splice list were each of type \li{Int}.

This parameterization strategy makes enforcing the binding discipline described in Sec.~\ref{sec:hygiene}
straightforward. 
Context independence is maintained by allowing the parameterized expansion 
to depend only on bindings given in the explicit \li{context}, which 
the system maintains bound relative to the definition site. 
The splices can 
depend only on the client site typing context. The use of function application ensures
that splices are capture avoiding. 
% We consider this more formally next.

Note that Hazel does \emph{not} statically check the definition of \li{expand}
to ensure that the encoded parameterized expansion has the necessary type in the specified \li{context}. Instead, the parameterized expansion 
is only \emph{validated} at each livelit invocation site, with errors reported to the client. 
A \emph{typed quotation} system as in, e.g., MetaOCaml \cite{DBLP:conf/flops/Kiselyov14}, could be adapted to allow for definition-site verification, as discussed in \cite{TLMs}, but 
note that   
the type of the quotation depends on the type of each splice in 
the splice list, which has arbitrary length in general (e.g. the 
number of splices in the \li{\$dataframe} livelit 
depends on how many rows and columns the user has added).